---
title: "Seascape characterization of Tropical Eastern Pacific monthly data"
author: "Erick Bolaños Duran"
date: "`r Sys.Date()`"
output:
    rmdformats::readthedown:
    code_folding: show
    highlight: kate
    toc_depth: 2
    self_contained: true
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style>
  #content {
    max-width: 1400px;
  }
  #sidebar h2 {
    background-color: #008B8B;
  }
  
  h1, h2 {
    color: #008B8B;
  }
</style>
```
```{r setup, include=FALSE}


knitr::opts_chunk$set(warning=FALSE, message=FALSE, 
                      echo = TRUE,
                      fig.path="figures_Rtemplate/",
                      fig.width = 10,
                      fig.height = 10,
                      fig.align="center")
library(dplyr)
library(ggplot2)
library(plotly)
library(googleVis)
library(knitr)
library(kableExtra)
library(DT)
library(summarytools)
library(forcats)
library(ggpubr)

```

# Sea surface Salinity (SSS)of TEP from 1998 to 2020 (Monthly records)
<hr>
## Overal mean 
<p> Open ***.nc*** of monthly records that was retrieved from Copernicus satellite images.Then, the overall mean  and SD will be calculated from the raster objects.</p>
```{r OM SSS, warning=FALSE, message=FALSE,fig.align="center", fig.width=10, fig.height=10, fig.cap="Fig.1 SSS Overall mean of TEP from  1998 to 2020" }
## ==============================================================
  
  # Title:2_Monthly_variables preparation
  # Author details: Erick Bolaños Duran
  # Contact details: ebduran@pg.cibnor.mx
  # Script and data info: This script help to prepare all raster images with a monthly time scale
  # Copyright (c): 2023 Erick Bolaños Duran

## ==============================================================

  
## Dependencies ==============================================================
library(rerddapXtracto)
library(lubridate)
library(ncdf4)
library(httr)
library(rgdal)
library(spatialEco)
library(raster)
library(dplyr)
library(tidyr)
library(lubridate)
library(data.table)
library(rgeos)
library(sp)
library(usdm)
library(raster)
library(sf)
library(udunits2)
library(rnaturalearth) 
library(rnaturalearthdata) 
##Set directory 
#Open monthly record of SSS
setwd("~/Doctorado/Acanthaster/Seascape genomics data/01_Seascape_characterization/02_SSS_POT/")
SSS_POT=brick("SSS_POT_1998-2020.nc")# The monthly data of the POT begin in January 1998 until December 2019
SSS_POT
#Calculate de overall mean
SSS_OM = mean(SSS_POT, na.rm=T)
plot(SSS_OM)
setwd("C:/Users/Erick Bolaños Durán/Documents/Doctorado/CIBNOR/fourth semester/01_Paper_submition/02_Data/02_Monthly_data/01_SSS_data")
writeRaster(SSS_OM,"OM_SSS_POT_1998-2020.tif",overwrite=TRUE)
```
## Overal SD 
```{r OSD SSS, warning=FALSE, message=FALSE,fig.align="center", fig.width=10, fig.height=10,fig.cap="Fig.2 SSS Overall SD of TEP from  1998 to 2020"}
##Calculate de overall DSD
SSM_OSD = calc(SSS_POT, sd, na.rm=T)
plot(SSM_OSD)
setwd("C:/Users/Erick Bolaños Durán/Documents/Doctorado/CIBNOR/fourth semester/01_Paper_submition/02_Data/02_Monthly_data/01_SSS_data")
writeRaster(SSM_OSD,"OSD_SSS_POT_1998-2020.tif",overwrite=TRUE)
```
# Attenuation Coefficient (Kd490) 
<p>First, Kd490 monthly records were downloaded from Copernicus t
##Overall mean
```{r OM KD, warning=FALSE, message=FALSE,fig.align="center", fig.width=10, fig.height=10, fig.cap="Fig.3 Kd490 Overall mean of TEP from  1998 to 2020"}
#Open monthly record of SSS 
setwd("~/Doctorado/Acanthaster/Seascape genomics data/01_Seascape_characterization/05_KD49_POT/")
kd49_POT=brick("kd490_POT_1998-2020.nc")
kd49_POT
#Calculate de overall mean
kd49_OM = mean(kd49_POT, na.rm=T)
plot(kd49_OM)
setwd("C:/Users/Erick Bolaños Durán/Documents/Doctorado/CIBNOR/fourth semester/01_Paper_submition/02_Data/02_Monthly_data/02_Kd490_data")
writeRaster(kd49_OM,"OM_kd490_POT_1998-2020.tif",overwrite=TRUE)
```
## Overal SD 
```{r OSD KD, warning=FALSE, message=FALSE,fig.align="center", fig.width=10, fig.height=10, fig.cap="Fig.4 Kd480 Overall SD of TEP from  1998 to 2020"}
##Calculate de overall SD
kd490_OSD = calc(kd49_POT, sd, na.rm=T)
plot(kd490_OSD)
setwd("C:/Users/Erick Bolaños Durán/Documents/Doctorado/CIBNOR/fourth semester/01_Paper_submition/02_Data/02_Monthly_data/02_Kd490_data")
writeRaster(kd490_OSD,"OSD_K490_POT_2003-2020.tif",overwrite=TRUE)
```
# Chlorophyll a (Chla) concentration of TEP from 1998 to 2020 (Monthly records)
## Overall mean
<p>
```{r OM CHA, warning=FALSE, message=FALSE,fig.align="center", fig.width=10, fig.height=10, fig.cap="Fig.5 Chlorophyll a Overall mean of TEP from  1998 to 2020"}
setwd("~/Doctorado/Acanthaster/Seascape genomics data/01_Seascape_characterization/03_Chla_POT/")
cha_98_99=brick("Chla_POT_1998-1999.nc")
cha_00_20=brick("Chla_POT_2000_2020.nc")
cha_POT=stack(cha_98_99,cha_00_20)
cha_POT
writeRaster(cha_POT,"cha_POT_1998-2020.tif", overwrite=TRUE)
#Calculate de overall mean
cha_OM = mean(cha_POT, na.rm=T)
plot(cha_OM)
setwd("C:/Users/Erick Bolaños Durán/Documents/Doctorado/CIBNOR/fourth semester/01_Paper_submition/02_Data/02_Monthly_data/03_Cha_data")
writeRaster(cha_OM,"OM_cha_POT_1998-2020.tif",overwrite=TRUE)
```
## Overall SD
```{r OSD CHA, fig.align="center", fig.cap="Fig 6. Chlorophyll a Overall SD of TEP from  1998 to 2020",fig.height=10, fig.width=10}
cha_OSD = calc(cha_POT, sd, na.rm=T)
plot(cha_OSD)
setwd("C:/Users/Erick Bolaños Durán/Documents/Doctorado/CIBNOR/fourth semester/01_Paper_submition/02_Data/02_Monthly_data/03_Cha_data")
writeRaster(cha_OSD,"OSD_cha_POT_2003-2020.tif",overwrite=TRUE)
```
# pH of TEP from 1998 to 2020 (Monthly records)
## Overall mean
```{r OM PH, warning=FALSE, message=FALSE,fig.align="center", fig.width=10, fig.height=10, fig.cap="Fig.7 pH Overall mean of TEP from  1998 to 2020"}
setwd("~/Doctorado/Acanthaster/Seascape genomics data/01_Seascape_characterization/04_Other_characterization_POT/")
pH_POT=stack("ph_POT_1998_2020.nc")
pH_POT
#Calculate de overall mean
pH_OM = mean(pH_POT, na.rm=T)
plot(pH_OM)
setwd("C:/Users/Erick Bolaños Durán/Documents/Doctorado/CIBNOR/fourth semester/01_Paper_submition/02_Data/02_Monthly_data/04_pH_data")
writeRaster(pH_OM,"OM_pH_POT_1998-2020.tif",overwrite=TRUE)
```
## Overall SD
```{r OSD PH, warning=FALSE, message=FALSE,fig.align="center", fig.width=10, fig.height=10, fig.cap="Fig.8 pH Overall SD of TEP from  1998 to 2020"}
pH_OSD = calc(pH_POT, sd, na.rm=T)
plot(pH_OSD)
setwd("C:/Users/Erick Bolaños Durán/Documents/Doctorado/CIBNOR/fourth semester/01_Paper_submition/02_Data/02_Monthly_data/04_pH_data")
writeRaster(pH_OSD,"OSD_pH_POT_2003-2020.tif",overwrite=TRUE)
```

# Process Sea Current Velocity from (Monthly records)
<p>For sea current velocity there is a preliminary step to take into account before calculating mean and standard deviation.CMEMS provides information on sea current velocity in two separate datasets: eastward velocity and northward velocity.</p> 
<hr>
<p> Since we are interested in the overall water velocity, we need to combine eastern and northern velocity. This can be done by considering eastern and northern velocity as perpendicular vectors of water velocity, and the euclidean norm of these vectors is the overall water velocity. 
##load raster - monthly resolution
```{r EWV and NWV, warning=FALSE, message=FALSE}
setwd("~/Doctorado/Acanthaster/Seascape genomics data/01_Seascape_characterization/9_Corrientes_velocidad/")
POT_EWV = stack('Westward_TEP_1998_2019.nc')
POT_NWV = stack("Northward_TEP_1998_2019.nc")
```
<p>In practice, the overall water velocity is the square root of the sum of squares of northward and eastward velocity</p>
```{r SCV, warning=FALSE, message=FALSE,fig.align="center", fig.width=10, fig.height=10}
POT_SCV = sqrt(POT_EWV^2+POT_NWV^2)
```
## Overall mean
```{r OM SCV, warning=FALSE, message=FALSE,fig.align="center", fig.width=10, fig.height=10, fig.cap="Fig.9 overall water velocity Mean"}
#Calculate de overall mean
POTSCV_OM = mean(POT_SCV, na.rm=T)
plot(POTSCV_OM)
setwd("C:/Users/Erick Bolaños Durán/Documents/Doctorado/CIBNOR/fourth semester/01_Paper_submition/02_Data/02_Monthly_data/05_SCV_data")
writeRaster(POTSCV_OM, 'POT_SCV_OM.tif', overwrite=T)
```
## Overall SD
```{r OSD SCV, warning=FALSE, message=FALSE,fig.align="center", fig.width=10, fig.height=10, fig.cap="Fig.10 Overall water velocity SD"}
POTSCV_OSD = calc(POT_SCV, sd, na.rm=T)
plot(POTSCV_OSD)
setwd("C:/Users/Erick Bolaños Durán/Documents/Doctorado/CIBNOR/fourth semester/01_Paper_submition/02_Data/02_Monthly_data/05_SCV_data")
writeRaster(POTSCV_OSD, 'POT_SCV_OSD.tif', overwrite=T)
```