---
title: "Selecting the best classification of ETP"
author: "Erick Bolaños Duran <p>***Email:*** ebduran@pg.cibnor.mx<br>"
date: "`r Sys.Date()`"
output:
    rmdformats::readthedown:
    code_folding: show
    highlight: kate
    toc_depth: 2
    number_sections: no
    self_contained: true
editor_options: 
  markdown: 
    wrap: 72
---
```{=html}
<style>
  #content {
    max-width: 1400px;
  }
  #sidebar h2 {
    background-color: #008B8B;
  }
  
  h1, h2 {
    color: #008B8B;
  }
</style>
```
```{r setup, include=FALSE}

#No MOVER!!!!!

knitr::opts_chunk$set(warning=FALSE, message=FALSE, #error=FALSE,
                      echo = TRUE,
                      fig.path="figures_Rtemplate/",
                      #dev='pdf', #Activar esta opción si se quiere obtener PDF
                      fig.width = 8,
                      fig.height = 4,
                      fig.align="center")
library(dplyr)
library(ggplot2)
library(plotly)
library(knitr)
library(summarytools)
library(forcats)
library(ggpubr)
library(kableExtra)


#Ubicación guia
#E:\X\Cursos\Taller Bioinformatica
```
# Selecting the best classification of POT
## Convert the SOM-PAM cluster into FCMres
<p>We can now compare the clustering analysis of the SOM-PAM classification and the FCM analysis using the SI and EI. For that, we need to convert the SOM-PAM in a FCMres objet.</p>
<p>To Create a FCMres</p>
<p>The parameters needed to create FCMres object</p>
<p>1.-Centers: a matrix representing the center of each group</p>
```{r getting the centers, warning=FALSE, message=FALSE}

## ==============================================================
  
  # Title:07_Clutering_comparation
  # Author details: Erick Bolaños Duran
  # Contact details: ebduran@pg.cibnor.mx
  # Script and data info: This script help to make a comparation among cluster analysis
  # Copyright (c): 2023 Erick Bolaños Duran

## ==============================================================
# Funtions===================================================================
### A funtion to Scale all variables to give them equal importance during the SOM training process. 
normalize =function(x) {
  min=raster::minValue(x)
  max=raster::maxValue(x)
  return((x - min) / (max - min))
}
###A funtion for color plots
coolBlueHotRed= function(n,alpha=1)
{rainbow(n, end = 4/6, alpha = alpha)[n:1]
  
}
## Dependencies ==============================================================
library(terra)
library(geocmeans)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(viridis)
library(spdep)
library(tmap)
library(cowplot)
library(geojsonio)
library(sf)
library(raster)
library(broom)
library(factoextra)
library(heatmaply)
library(plotly)
library(rassta)
library(kohonen)
######Open non correlated data
ft =list.files(path = "d:/Estudiantes/Erick_Bolaños/01_Tesis/01_Paper_submition/02_Data/05_Non_correlated_data/" ,full.names = T,pattern = ".tif$",all.files=TRUE)
rast_1=stack(ft)
#change names of layer
layer.names=c("MaxMMSST","minMMSST","OMKD490","OMpH","OM_SCV","OM_SSS","OSD_pH")
names(rast_1)=layer.names
###Scale the env variables using the previous function
rast_1=normalize(rast_1)
########They are already normalized, now we are going to open the raster with the terra package, for this we have to change the object with the rast function
env.data=terra::rast(rast_1)
# giving a name to each band
dataset <- lapply(names(env.data), function(n){
  aband <- env.data[[n]]
  return(aband)
})

###make the SOM and PAM
#############Grid 8x9#####
set.seed(963) ##to reproductibility 
tsom8x9<-som_gap(env.data, xdim =8, ydim=9,rlen = 1000, mode = "online", K.max =25 ,B = 1000, topo="hexagonal",
                 neighbourhood.fct = "gaussian", method = "globalSEmax")
####Make the PAM##
tpam8x9=som_pam(ref.rast= env.data, kohsom = tsom8x9$SOM, k=tsom8x9$Kopt)
#1.- getting the groups
codesSOM8x9=getCodes(tsom8x9$SOM)
codesSOM8x9=as.data.frame(codesSOM8x9)
class(codesSOM8x9)#To see what type of object it is, in this case a data.frame
####W have to add the PAM groups to a new collum in the data frame 
pamgap8x9=tpam8x9$sompam$clustering
class(pamgap8x9)##check the data class 
codesSOM8x9$pam <- as.integer(pamgap8x9)#Add the group number of PAM clustering
#In this case, we calculate the centers of the groups as the mean of each variable in each group.
centersPAM8x9 <- codesSOM8x9 %>% 
  group_by(codesSOM8x9$pam) %>%
  summarise_all(mean)
centersPAM8x9 <- as.data.frame(centersPAM8x9[2:9])  
```
<p>2.-Belongings: a membership matrix of each observation to each group</p>
```{r getting the belongings, warning=FALSE, message=FALSE}
#2.-The membership matrix is a simple binary matrix.
member_matPAM8x9 <- cat_to_belongings(codesSOM8x9$pam)
```
<p>3.-Data: the dataset used for the clustering in this case the CODEBOOKS derived from SOM</p>
```{r getting the data, warning=FALSE, message=FALSE}
#3.-The data
#Data: the data set used for the clustering in this case the CODEBOOKS derived from SOM
codesSOM8x9$pam  <- NULL
```
<p>4.-m: the fuzzyness factor (1 if using a hard clustering method) in this case PAM ins a hard clusterin method</p>
```{r Making the FCMres object, warning=FALSE, message=FALSE}

  #m: the fuzziness factor (1 if using a hard clustering method), in this case PAM ins a hard clusterin method
PAM_SOM8x9<- FCMres(list(
  "Centers" = centersPAM8x9,
  "Belongings" = member_matPAM8x9,
  "Data" = codesSOM8x9,
  "m" = 1,
  "algo" = "pam"
))
```
## Comparing quality metrics 
<p>Now, we can calculate the SI of the SOM-PAM using the FCMres object and the funtion ***calcSilhouetteIdx***</p>
```{r SI, warning=FALSE, message=FALSE}
#silhouette index
SI_SOMPAM8x9=calcSilhouetteIdx(PAM_SOM8x9$Data, PAM_SOM8x9$Belongings)
SI_SOMPAM8x9
#Registered S3 method overwritten by 'fclust':
#method       from 
#print.fclust e1071
#[1] 0.382964
```
<p>Now, we can calculate the SI of the SOM-PAM using the FCMres object and the funtion ***calcexplainedInertia***</p>
```{r EI, warning=FALSE, message=FALSE}
EI_SOMPAM8x9=calcexplainedInertia(PAM_SOM8x9$Data, PAM_SOM8x9$Belongings)
EI_SOMPAM8x9
#[1] 0.8990287
###ADD SOMPAM quality index into indexFCM
```
<p>Now, we can create a table with all the clustering quality indexes of all the ETP classification methods</p>
```{r making a table, warning=FALSE, message=FALSE}
#FCM
FCM_result =CMeans(dataset, k = 13, m = 1.2, standardize = F,
                  verbose = FALSE, seed = 789, tol = 0.001, init = "kpp")
#GFCM
GFCM_result <- GCMeans(dataset, k = 13, m = 1.2, beta = 0.4, standardize = F, verbose = FALSE, seed = 789, tol = 0.001)
#SGFCM
w1 <- matrix(1, nrow = 3, ncol = 3)
SGFCM_result <- SGFCMeans(dataset, k = 13, m = 1.2, standardize = F,
                          lag_method = "mean",
                          window = w1, alpha = 0.5, beta = 0.4,
                          seed = 789, tol = 0.001, verbose = FALSE, init = "kpp")

### To obtain the EI and SI of FCM results
cluster_results <- list(FCM_result, GFCM_result, SGFCM_result)
indexFCM <- sapply(cluster_results, function(clust){
  c(calcexplainedInertia(clust$Data, clust$Belongings),
    calcSilhouetteIdx(clust$Data, clust$Belongings))
})
colnames(indexFCM) <- c("FCM", "GFCM", "SGFCM")
rownames(indexFCM) <- c("explained inertia", "silhouette index")
### To add the EI and SI of SOM-PAM un the FCM results
SOMPAM8x9=c(EI_SOMPAM8x9,SI_SOMPAM8x9)
Qindex2 <-cbind(indexFCM[, c(1,2,3)],SOMPAM8x9)


QindexR2=knitr::kable(Qindex2, digits = 3)
QindexR2
```
<p>We can create a heatmap with all the clustering quality indexes of all the ETP classification methods</p>
```{r making a heatmap, warning=FALSE, message=FALSE}
####Quality clustering index  comparative using a Heat map
heatmap <- heatmaply(Qindex2,
                     dendrogram = "both",
                     xlab = "Quality clustering index", ylab = "Clustering methods", 
                     main = "Comparative among clustering methods",
                     scale = "none",
                     #margins = c(60,100,40,20),
                     grid_color = "white",
                     grid_width = 0.00001,
                     hide_colorbar = FALSE,
                     branches_lwd = 0.5,
                     fontsize_row = 12, fontsize_col = 12,
                     labCol = colnames(Qindex2),
                     labRow = rownames(Qindex2),
                     label_names = c("Modelo", "Metric:", "Value"),
                     heatmap_layers = theme(axis.line=element_blank()))

heatmap 
```
